Quantifying Spatially Resolved Hydration Thermodynamics Using Grid Inhomogeneous Solvation Theory [Article v\versionnumber]

### 2022-12-15 manuscript/manuscript.tex
Small changes

\subsubsection{MD simulation}
Now, you can run restrained MD simulations based on the equilibrated structures.
Restraints should be used because IST theory is not defined for moving solutes.
We apply a harmonic restraint of \SI{100}{\kilo\calorie\per\mole\per\angstrom\squared} to all heavy atoms to keep the system in place.
We run \SI{100}{\nano\second} of simulation and store the conformations every \SI{100}{\pico\second}.
A simulation length of at least \SI{20}{\nano\second} is recommended to obtain statistically independent snapshots of the water molecules.

An example Amber input file for GIST might look like this:
\begin{lstlisting}[style=amber-in]
restrained 100 ns NpT
&cntrl
ntx=5, irest=1,
ioutfm=1,
ntb=2, iwrap=1,
ntr=1, restraint_wt=100.0,
restraintmask="!@H=&!:WAT",
ntp=1, pres0=1.0, taup=1.0,
ntc=2, ntf=2,
ntt=3, tempi=300.0, temp0=300.0, gamma_ln=2,
nstlim=50000000, dt=0.002,
ntwr=50000, ntwx=5000, ntpr=5000,
/
\end{lstlisting}
Note that these are very standard MD settings except for the restraints.

Run the MD using \software{pmemd.cuda}:
\begin{lstlisting}[style=bash]
pmemd.cuda \
-O \
-i 100ns-npt-restraint.in \
-o md-01.out \
-p solvated.parm7 \
-c EQUIL-DONE.rst \
-x md-01.nc \
-r md-01.ncrst \
-ref EQUIL-DONE.rst
\end{lstlisting}

\subsection{Applying GIST for the Streptavidin-Biotin complex}
\subsubsection{System Preparation and Equilibration}
If you want to skip this section, run 
\begin{lstlisting}[style=bash]
make equilibration-targets
\end{lstlisting}
in the code folder. This assumes that you have Python and Amber set up properly and that \software{pmemd.cuda} is in the PATH.

The choice of initial structure is crucial to the GIST results.
Crystal structures may not represent the dominant solution structure for a given target.
In our example, we only use a single equilibrated structure.
A more sophisticated approach would be to identify one or several dominant structures out of the ensemble.
Depending on what properties you are interested in, it might be beneficial to run an MD simulation choose dominant structures through clustering.

When starting from a crystal structure, care has to be taken during the preparation.
Make sure to choose a proper protonation state for all titratable moeties.
You will need to solvate your structure in a solvent box, which should be large enough to obtain unbiased long-range electrostatics.
A good criterion for validating the box size is to check whether the water on the outside of the box behaves bulk-like according to the GIST output.

@@

Use the \software{equilibration.py} script (or your own) to perform short NVT and NpT equilibration runs on the complex structure.
It is important to include unrestrained NpT steps, because restraining two molecules at the same time might interfere with the pressure equilibration.

>>>>

\subsubsection{MD simulation}
\todo{Move simulation part from complex to here}
Now, you can run restrained MD simulations based on the equilibrated structures for subsequent GIST analyses.
Restraints should be used during the MD because IST theory is not defined for moving solutes.
We apply a harmonic restraint of \SI{100}{\kilo\calorie\per\mole\per\angstrom\squared} to all heavy atoms to keep the system in place.
We run \SI{100}{\nano\second} of simulation and store the conformations every \SI{100}{\pico\second}.
A simulation length of at least \SI{20}{\nano\second} is recommended to obtain statistically independent snapshots of the water molecules.

An example Amber input file for GIST might look like this:
\begin{lstlisting}[style=amber-in]
	restrained 100 ns NpT
	&cntrl
	ntx=5, irest=1,
	ioutfm=1,
	ntb=2, iwrap=1,
	ntr=1, restraint_wt=100.0,
	restraintmask="!@H=&!:WAT",
	ntp=1, pres0=1.0, taup=1.0,
	ntc=2, ntf=2,
	ntt=3, tempi=300.0, temp0=300.0, gamma_ln=2,
	nstlim=50000000, dt=0.002,
	ntwr=50000, ntwx=5000, ntpr=5000,
	/
\end{lstlisting}
This file will run a 100 ns NpT MD simulation while keeping all atoms except for hydrogens and waters restrained. Note that these are very standard MD settings except for the restraints.

Run the MD using \software{pmemd.cuda}:
\begin{lstlisting}[style=bash]
	pmemd.cuda \
	-O \
	-i 100ns-npt-restraint.in \
	-o md-01.out \
	-p solvated.parm7 \
	-c EQUIL-DONE.rst \
	-x md-01.nc \
	-r md-01.ncrst \
	-ref EQUIL-DONE.rst
\end{lstlisting}
\subsubsection{GIST analysis}
To start our GIST analysis of biotin, we'll first have to decide on the parameters defining the GIST grid during the analyis.
We recommend that you center the biotin molecule in space:
\begin{lstlisting}[style=cpptraj]
	parm solvated.parm7
	trajin md-01.nc 1 1 1
	autoimage !(:WAT) origin
	strip :WAT
	trajout biotin-centered.pdb
\end{lstlisting}

To decide on a grid size, find the extent (minimum and maximum) of the x-, \mbox{y-,} and z-coordinates of the molecule.
This can be done e.g., using the \software{mdtraj} package in Python.
\begin{lstlisting}[style=python]
	import mdtraj as md
	pdb = md.load("biotin-centered.pdb")
	# mdtraj converts to nm.
	# *10 to convert back to Angstrom.
	min = pdb.xyz[0].min(0) * 10.
	max = pdb.xyz[0].max(0) * 10.
	print(*list(zip("xyz", min, max)), sep="\n")
\end{lstlisting}
For a system centered at the origin with a spacing of \SI{0.5}{\angstrom} and a buffer of, e.g., \SI{15}{\angstrom} between the solute and the grid wall, you need $4.0\times (d+15)$ bins in the respective direction, where $d$ is the larger of $-$min and $+$max in the respective direction.
To get bulk-like water properties at the grid border, a buffer distance of \SI{15.0}{\angstrom} is reasonable for biotin.

Now, you can finally run the actual GIST analysis.
Center the molecule as above, to make sure that its position matches the grid parameters you decided on.
We recommend using either the PME or GPU implementation of GIST.
For the GPU version, cpptraj needs to be compiled with the \inlinecode{-cuda} flag.
While the GPU version is generally faster, the PME implementation matches the energy of the Amber MD engine more closely.

\begin{lstlisting}[style=cpptraj]
	parm solvated.parm7
	trajin md-01.nc
	autoimage !(:WAT) origin
	# optionally remove the "pme" option
	gist griddim 100 100 100 out gist.dat \
	refdens 0.03287 pme
\end{lstlisting}
\subsubsection{Interpreting and visualizing the results}
\todo{write this, show the various GIST properties here: Eww, Esw, S. Show how to calculate A and E with gisttools}
\subsection{Applying GIST for the Streptavidin-Biotin complex}
\subsubsection{System Preparation and Equilibration}
If you want to skip this section, run 
\begin{lstlisting}[style=bash]
make equilibration-targets
\end{lstlisting}
in the code folder. This assumes that you have Python and Amber set up properly and that \software{pmemd.cuda} is in the PATH.

The choice of initial structure is crucial to the GIST results.
Crystal structures may not represent the dominant solution structure for a given target.
In our example, we only use a single equilibrated structure.
A more sophisticated approach would identify one or several dominant structures out of the ensemble - either gathered from experimental data or by clustering an MD simulation.

When starting from an experimental structure, care has to be taken during the preparation step.
Make sure to choose a proper protonation state for all titratable moeties.
You will need to solvate your structure in a solvent box, which should be large enough to obtain unbiased long-range electrostatics.
A good criterion for validating the box size is to check whether the water on the outside of the box behaves bulk-like according to the GIST output.

@@

Use the \software{equilibration.py} script (or your own procedure) to perform short NVT and NpT equilibration runs on the complex structure.
Note that the equilibration script contains unrestrained NpT steps, because restraining two molecules at the same time might interfere with the pressure equilibration.
